---
title: "The Linguistic Landscape of Stream-of-Consciousness Literature"
subtitle: "Exploring Word Frequency and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned, transnational stream of consciousness authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By analyzing the word frequency of their famous works, I explore mental health themes like anxiety, depression, trauma, and existential angst. Through word frequency analysis of nine novels, I aim to uncover shared linguistic patterns and gain insights into the authors' mental states, offering a glimpse into themes of self-identity and existential contemplation."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tidytext)
library(rmarkdown)
library(gutenbergr)
library(tm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(igraph)
library(widyr)
library(ggraph)

portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))

```


# Introduction

Stream of consciousness is a narrative technique that aims to capture the continuous flow of thoughts, feelings, and sensations experienced by a character without conventional organization or punctuation  [@resampling]. It mirrors the unpredictable and interconnected nature of human thought processes, often revealing the inner workings of the character's mind in an intimate and unfiltered manner [@flow]. In literature, most scholars agree that stream of consciousness reveals the complexities of mental-scapes, shedding light on the nuances of characters' emotional well-being and psychological struggles [@pathology]. Its relationship to mental health lies in its ability to offer insight into the depths of consciousness, providing a platform to explore themes such as anxiety, trauma, and existential questioning. As such, this paper has mined the texts of a total of nine novels from the volunteer archive, Project Gutenberg, to examine the mental health themes of famous stream of consciousness authors, namely by Joyce, Woolf, Proust, Mansfield and Eliot, from the modernist era of literature, spanning from late 19th century to the mid-20th century [@gutenberg]. (more stats and data mentioned here)

By analyzing these textual datasets, I seek to pose and answer crucial questions: What are some important factors contributing to this relationship between mental health, disassociation and stream of consciousness? Moreover, how does this relationship vary differently across differnet demographics of authors, for instance, authors with different geographical locations and genders? Understanding these dynamics is crucial in having an informed understanding of the West's late 19th to mid-20th century socio-political landscape, especially in regards to how authors and creative writers navigate and deal with then-taboo topics such as existential angst, mental health issues and disabilities.

Thus, my estimand is the correlation between famous stream of consciousness novels, words related to mental health themes and their frequency. This is considered in terms of the nine selected novels only, namely Joyce's A Portrait of the Artist as a Young Man and Chamber Music; Woolf's Mrs Dalloway and Jacob's Room; Proust's Swann Way; Mansfield's Bliss and The Garden Party; and Eliot's The Waste Land and The Love Song of J. Alfred Prufrock. Through our analysis, we found that (percentage, number and data here, main results)...

To further understand the correlation between stream of consciousness novels and mental health themes, in [Introduction], our paper briefly discusses the nature of stream of consciousness literature, relevant authors and the works that we've chosen to analyze. Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts. Next, [Discussion] provides further insights and future areas of study. Finally, [Conclusion] summarizes our main findings. To complete the paper, [Appendix] clarifies how each variable within each dataset is generated and tables to accordingly demonstrate this.

The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg] `tm` [@rTm] and `knitr` [@rKnitr].

# Data {#sec-data}

## Measurement

## Source Data

## Data Cleaning and Word Tokenization

```{r}
#| echo: false
#| message: false
#| label: tbl-genderclass
#| tbl-cap: Table of Number of Classes Students Considered for Regrade Requests by Students’ Gender

```

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

### Word Count

```{r}
#| echo: false
#| eval: true
#| label: tbl-joyce
#| tbl-cap: "Examplary Table of Unprocessed Stream of Consciousness Novels (James Joyce)"
#| warning: false

portrait |>
  rename("Book ID" = gutenberg_id,
         "Text" = text,
         "Book" = book) |>
  head(10) |>
  kable(
    col.names = c("Book ID", "Text", "Book"),
    booktabs = TRUE
  )

```


```{r}
#| echo: false
#| eval: true
#| label: tbl-joyce-token
#| tbl-cap: "Examplary Table of Tokenized Stream of Consciousness Novels (James Joyce)"
#| warning: false

portrait_clean |>
  rename("Book ID" = gutenberg_id,
         "Book" = book,
         "Word" = word) |>
  tail(10) |>
  kable(
    col.names = c("Book ID", "Book", "Word"),
    booktabs = TRUE
  )

```


```{r}
#| echo: false
#| eval: true
#| label: tbl-joyce-words
#| tbl-cap: "Examplary Table of Word Count of Stream of Consciousness Novels (James Joyce)"
#| warning: false

portrait_words |>
  rename("Word" = word) |>
  head(10) |>
  kable(
    col.names = c("Word", "Count"),
    booktabs = TRUE
  )

```

### Comparative Word Frequency

```{r}
#| echo: false
#| eval: true
#| label: tbl-frequency
#| tbl-cap: "Word Frequency of Stream of Consciousness Novels, A Comparison Between Five Authors "
#| warning: false

frequency <- bind_rows(mutate(portrait_words, author = "James Joyce"),
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(dalloway_words, author = "Virginia Woolf"),
                       mutate(bliss_words, author = "Katherine Mansfield"),
                       mutate(prufrock_words, author = "T.S. Eliot")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) 

frequency |>
  rename("Word" = word) |>
  head(12) |>
  kable(
    col.names = c("Word", "James Joyce", "Katherine Mansfield", "Marcel Proust", "T.S. Eliot", "Virignia Woolf"),
    booktabs = TRUE
  )

```

### Generating Word Networks

# Model

## Model set-up

### Model justification


# Results

## The Dominant Vocabulary of Stream of Consciousness

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-count
#| layout-ncol: 2
#| fig-cap: "Comparative Analysis of Top 20 Word Frequencies by Stream of Consciousness Authors' Novels"
#| fig-subcap: ["James Joyce", "Virginia Woolf", "Marcel Proust", "T.S. Eliot", "Katherine Mansfield"]

portrait_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(x = "Frequency", y = "Word")

dalloway_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(x = "Frequency", y = "Word")

swann_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(x = "Frequency", y = "Word")

prufrock_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(x = "Frequency", y = "Word")

bliss_words |> head(20) |>
ggplot(aes(x = n, y = reorder(word, n))) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(x = "Frequency", y = "Word") 


```

## Great Britain as the Center of Stream of Consciousness Literature

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-frequency-brit
#| fig-cap: "Comparative Analysis of Word Frequency in British Stream of Consciousness Authors"

frequency_britain <- bind_rows(
                       mutate(dalloway_words, author = "Virginia Woolf"),
                       mutate(prufrock_words, author = "T.S. Eliot"),
                       mutate(portrait_words, author = "James Joyce")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`T.S. Eliot`:`James Joyce`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_britain, aes(x = proportion, y = `Virginia Woolf`, color = abs(`Virginia Woolf` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.05, size = 1.5, width = 0.7, height = 0.7) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslateblue", high = "gray1") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "Virginia Woolf", x = NULL)

```

## Gendered Mental Landscape of Stream of Consciousness Novels

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-gender
#| layout-ncol: 2
#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male Stream of Consciousness Authors"
#| fig-subcap: ["Female Authors", "Male Authors"]


# Calculate the word frequency over all unique words and bind all frequency rate of authors together to conduct a comparative comparison
# Code sourced from https://medium.com/the-data-nudge/nlp-basics-exploring-word-frequency-with-the-tidytext-r-package-ac35a6d805f4

  
frequency_female <- bind_rows(mutate(dalloway_words, author = "Virginia Woolf"),
                              mutate(bliss_words, author = "Katherine Mansfield")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Virginia Woolf`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_female, aes(x = proportion, y = `Katherine Mansfield`, color = abs(`Katherine Mansfield` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1, width = 0.5, height = 0.5) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = scales::percent_format()) +
  scale_y_log10(labels = scales::percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslateblue", high = "gray50") +
  facet_wrap(~author, ncol = 1) +
  theme(legend.position="none") +
  labs(y = "Katherine Mansfield", x = NULL)

frequency_male <- bind_rows(
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(prufrock_words, author = "T.S. Eliot"),
                       mutate(portrait_words, author = "James Joyce")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Marcel Proust`:`James Joyce`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_male, aes(x = proportion, y = `T.S. Eliot`, color = abs(`T.S. Eliot` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1.5, width = 0.8, height = 0.8) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslateblue", high = "gray100") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "T.S. Eliot", x = NULL)


```

### Comparing Female Stream of Consciousness Authors

### Comparing Male Stream of Consciousness Authors


## Transnational Stream of Consciousness Novels and Mental Health Themes

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-frequency-int
#| fig-cap: "Comparative Analysis of Word Frequency in Transnational Stream of Consciousness Authors"

frequency_international <- bind_rows(
                       mutate(swann_words, author = "Marcel Proust"),
                       mutate(bliss_words, author = "Katherine Mansfield"),
                       mutate(portrait_words, author = "James Joyce")) |>
  mutate(word = str_extract(word, "[a-z']+")) |>
  count(author, word) |>
  group_by(author) |>
  mutate(proportion = n / sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = author, values_from = proportion) |>
  pivot_longer(`Katherine Mansfield`:`James Joyce`,
               names_to = "author", values_to = "proportion")

ggplot(frequency_international, aes(x = proportion, y = `Marcel Proust`, color = abs(`Marcel Proust` - proportion))) +
  geom_abline(color = "gray50", lty = 2) +
  geom_jitter(alpha = 0.05, size = 1, width = 0.8, height = 0.8) +  
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslateblue", high = "gray100") +
  facet_wrap(~author, ncol = 1) +
  theme(legend.position="none") +
  labs(y = "Marcel Proust", x = NULL)

```

## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false

## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o

generate_word_graph <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 200) 

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 0.2)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)

}


# Code sourced from
# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100

# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))

```


```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-networks
#| layout-ncol: 2
#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one

combined_text|>
  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)


combined_text|>
  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)

combined_text|>
  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
```
# Discussion

## Mental Health Vocabulary: Patterns and Trends 

Discuss vocabulary patterns and word trends

## Insights into Socio-Political Landscape of the West's Modernist Era 

The novels' linguistic patterns reflect the socio-political landscape of the Western hemisphere, from  late 19th century to the mid-20th century.

## Schizophrenic and Disassociative Tendencies in Female Stream of Consciousness


## Weaknesses 

### Lack of Thorough Word Cleaning

### Limiting Literature Works and Publication Editions

### Project Gutenberg's Focus on the Canon

## Moving Forward and Next Steps


\newpage

# Appendix

## Additional Data Details

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#combined_books 
```

### Data Gathering

```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-reasons-strip-search
#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches

  
```

### Data Cleaning


```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-items-strip-search
#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches


```


## Model Details {#sec-model-details}

## Posterior predictive check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics


```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2


```


\newpage

# References


